### 2017-12-18 Draft_Form/saerlp_article_master.tex
Initial commit 18_12_2017
Initital commit

\documentclass[a4paper, 12pt, titlepage]{scrartcl} % tech.report
\usepackage{authblk}

%============================================ Define Titlepage =====================================================%

\author[*$\dag$]{\large Andreas Hill}
\author[*]{\large Daniel Mandallaz}
\author[**]{\large Joachim Langshausen}
\affil[*]{\normalsize Department of Environmental Systems Science, ETH Zurich}
\affil[**]{\normalsize Forest Service Rhineland-Palatinate}       
\renewcommand\Authands{, }

% \date{\vspace{-0.3cm} \large \today \vspace{1.1cm}}
\date{\vspace{-0.8ex}}
\title{\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation of timber volume resources on forest district levels \\ \vspace{2ex}
  \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}}
\subject{\vspace{-1.5cm} Article Draft \vspace{0.2cm} \vspace{-0.5cm}}

% alternative title(s):
% - Implementation of design-based small area estimations on forest district level in Rhineland-Palatinate (Germany) by combining remote sensing information with data of the Third German National Forest Inventory
%
% Large scale implementation of design-based small area estimations on forest district level in Rhineland-Palatinate using data of the Third German National Forest Inventory
% Implementation of design-based small area estimations on forest district level in Rhineland-Paltinate (Germany) by
% combining remote sensing information with data of the Third German National Forest Inventory
%
% Large-scale implementation of design-based small area estimations on forest district level by
% combining remote sensing data with data of the Third German National Forest Inventory
%
% A double-sampling extension of the Third German National Forest Inventory for small area estimations of timber volume resources on forest district level with a large scale application to the federal state of Rhineland-Palatinate
%
%

%============================================ Load Packages ========================================================%
\usepackage{hyperref}
\usepackage{natbib}
 \hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    filecolor=magenta,      
    urlcolor=blue}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}%package zum einbinden von Bildern
\usepackage{fancyhdr}
\usepackage{amsmath} %Paket für erweiterte math. Formeln
\usepackage{amsfonts}
\usepackage{mathptmx}
\usepackage{booktabs}%Weitere Optionen für Tabellen
\usepackage{array} 
\usepackage{float}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{boldline}
\usepackage[explicit]{titlesec}
\usepackage{listings} %für R-Skripte
\usepackage{color}
\usepackage{pdfpages} % to include pdfs
\usepackage{graphicx}
\usepackage[font=footnotesize]{caption} % define font-size for captions
\usepackage[font=footnotesize]{subcaption}
\usepackage{rotating}
\usepackage{longtable} 
\usepackage{dcolumn} 
\usepackage{pictex}
\usepackage{tikz} 
 \usetikzlibrary{external}       % function to externalize tics-graphic buildings
 \tikzset{external/force remake} % overwrites existing externalized pdfs
 \tikzexternalize[prefix=cache/] % All externalized pdf are saved in folder 'cache'
 \tikzexternaldisable % Only externalise on-demand

% comment out to disable line-numbering:
%\usepackage{lineno}
%\linenumbers

%\bibliographystyle{ieeetr}
%\bibliographystyle{apasoft}
\bibliographystyle{spbasic}
%\bibliographystyle{spmpsci}

%============================================ Define Settings ========================================================%

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.2,0.7,0.25}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  %frame=single,                    % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Octave,                 % the language of the code
  morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text 
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,                       % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; 
                                   %  also try caption instead of title
}

\titlespacing{\section}{50pt}{2em}{2em} % {left}{before-sep}{after-sep}
\titlespacing{\subsection}{14pt}{2em}{1.5em}
\titlespacing{\subsubsection}{12pt}{2em}{1em}

%\setcounter{page}{1} % Setzt Seitenzahl gleich 1
\setlength{\parindent}{1em} % Einzug bei neuen Absätzen

%============================================ Define personal commands=============================================%

% remind myself of something:
\newcommand{\note}[1]{\color{red}(#1!)\color{black}}
\newcommand{\missref}{\note{[REF]}}
\newcommand{\todo}[1]{\textcolor{blue}{[Todo: #1]}}

% reocurring terms:
\newcommand{\adjrsq}{adjusted $R^2$}
\newcommand{\rmsecv}{RMSE$_{cv}$}
\newcommand{\mha}{m$^3$/ha}
\newcommand{\var}{\mathbb{V}}
\newcommand{\EX}{\mathbb{E}}
\newcommand{\psynth}{\textit{PSYNTH}}
\newcommand{\psmall}{\textit{PSMALL}}
\newcommand{\extpsynth}{\textit{EXTPSYNTH}}

%============================================ Start Document =======================================================%

\begin{document}\sloppy

%============================================ SetUp directories ====================================================%

\maketitle
\thispagestyle{empty}
\newpage

\vspace*{\fill}
\begin{tabular}{lll}

\small Corresponding author$^\dag$: & \small \underline{andreas.hill@usys.ethz.ch} \\
\small Co-authors: & \small \underline{daniel.mandallaz@env.ethz.ch}, \underline{joachim.langshausen@wald-rlp.de} \\
\end{tabular}\\

\begin{tabular}{lll}
\small Publication date: & \small \today \\
\end{tabular}
\newpage

%  Abstract
\input{tex/Abstract}\thispagestyle{empty} \newpage\cleardoublepage

% begin roman pagenumbering
\pagenumbering{Roman}

\tableofcontents
\newpage

\listoffigures
\newpage

\listoftables
\newpage

\pagestyle{fancy} %Kopfzeile und Fusszeile
\fancyfoot[C]{\thepage}
\setlength{\headsep}{15mm}

%============================================ Test temp ===============================================%

%\input{tex/test_temp}\newpage\cleardoublepage

%============================================ Introduction ========================================================%

\pagenumbering{arabic}
\setcounter{page}{1}

\input{tex/Introduction(2)}\newpage\cleardoublepage

%============================================ Materials and Methods ===============================================%

\input{tex/Materials_and_Methods(2)}\newpage\cleardoublepage

%============================================ Results =============================================================%

%\input{tex/Results}\newpage\cleardoublepage

%============================================ Discussion ==========================================================%

%\input{tex/Discussion}\newpage\cleardoublepage

%==================================== Conclusion & Outlook ==========================================================%

%\input{tex/Conclusions}\newpage\cleardoublepage

%============================================ Acknowledgements ==========================================================%

%\input{tex/Acknowledgements}\newpage\cleardoublepage

%============================================ Appendices ==========================================================%

%\input{tex/Appendices}\newpage\cleardoublepage

%============================================ Literature ==========================================================%

\bibliography{bib/literaturerRLPsae}\newpage\cleardoublepage

%==================================================================================================================%
\end{document}