### 2017-12-18 Draft_Form/tex/Acknowledgements.tex

\section*{Acknowledgements}
\thispagestyle{plain}

We want to express our gratitude to Prof. H. Heinimann (Chair of Land Use Engineering, ETH Zurich) for supporting this study.

We want to explicitly thank Dr. Johannes Stoffels and Dr. Henning Buddenbaum from the Environmental Sensing and Geoinformatic Group of University of Trier for providing the LiDAR data and tree species classification map.

Special gratitude is owed to the State Forest Service of Rhineland-Palatinate, in particular Jürgen Dietz and Claus-Andreas Lessander, for collaboration and providing the forest inventory- and geodata.

We also want to thank Kai Husmann and Christoph Fischer from the Northwest German Forest Research Institution Göttingen for their advice in processing the terrestrial inventory data.

% people to thank:
% - H.R. Heinimann
% - Johannes und Henning
% - Thomas Riedel (Thuenen-Institute) for providing the densified sample grid
% - Kai Husmann & Christoph Fischer