### 2017-12-18 Draft_Form/tex/Appendices.tex

%\appendix
%\addcontentsline{toc}{section}{Appendices}
%\section*{Appendices}
%\section{First appendix}
%\section{Second appendix}

\appendix
\section{Appendix}

% ----------------------------------------------------------------------- %
% A.1
% ----------------------------------------------------------------------- %
%\subsection{Model accuracies for $CHM$ variables}

%\input{tab/mod_chm_table.tex}\newpage