### 2017-12-18 Draft_Form/tex/Discussion.tex

\section{Discussion} %article
\label{sec:Dis}

%\subsection{Stratification according to Tree Species and LiDAR Acquisitions}
%\label{sec:strat_dis}

% NOTES:
%
% - Kuliesis et al.:  --> "NFI [provide reliable estimates for the most important attributes and] can be used 
%                          as a tool for validation of all other inventories", i.e. stand-wise inventories
%                     ==> in our case, we can say that we illustrated how the German NFI data can now be used
%                         for this purpose, i.e. providing of reliable data to validate the stand-wise inventories
%
%
% --> Extension to at least partially exhaustive estimator (three-phase non-exhaustive) if auxiliary data are really exhaustively
%     available
%
%
%
%
%
%
%
%
%
%