### 2018-02-10 Draft_Form/tex/Acknowledgements.tex
update 10_02_2018
update 10_02_2018

\section*{Acknowledgements}
\thispagestyle{plain}
We want to express our gratitude to Prof. H. Heinimann (Chair of Land Use Engineering, ETH Zurich) for supporting this study. We also want to explicitly thank Johannes Stoffels and Henning Buddenbaum from the Environmental Sensing and Geoinformatic Group of University of Trier for providing the ALS data and tree species classification map, and Kai Husmann and Christoph Fischer from the Northwest German Forest Research Institution Göttingen for their advice in processing the terrestrial inventory data. Special gratitude is also owed to Thomas Riedel from the Th{\"u}nen Institute for providing the densified NFI sample grid, and Alexander Massey for proofreading.